### 2015-04-21 OECD-TiVa-LMIC-GVCs.Rmd
add trca nrca plot

---
title: "OECD TiVa LMIC GVCs"
author: "Victor Kummritz and Bastiaan Quast"
date: "March 24, 2015"
output: ioslides_presentation
---

## Motivation

 - Global Value Chains relevance
 - GVCs in developement (capacity building)
 - General GVC picture overemphasises north-north because of higher monetary value

## This paper

## Data

ICIOs from TiVa (not WIOD)

TiVa data
 - 57 countries
 - 34 sectors
 - LMICs

## New Toolbox and Indicators

 - source / using countries specific indicators
 - decompor and gvc functions

 ```{r}
 library(decompr)
 ls.str("package:decompr")
 ```

 ```{r}
 library(gvc)
 ls.str("package:gvc")
 ```

 ```{r, echo=FALSE}
 # load the data
load("data/gvc_indicators.RData")
load("data/nrca_df.RData")
load("data/country_vars.RData")
load("data/w1995_2008.RData")
load("data/trca.RData")

 # load required libraries
library(dplyr)
library(ggvis)
library(magrittr)

# contruct avg_pop and gdp
country_vars %<>%
  group_by(ctry) %>%
  summarise(avg_pop = mean(pop)) %>%
  merge(country_vars, . , by = "ctry")

country_vars$avg_gdp <- with(country_vars, avg_pop * avg_gdppc)

# create factor gdp var
country_vars$ic      <- with(country_vars, ifelse(avg_gdppc <= 6000, "lic",
                                                  ifelse(avg_gdppc > 12000, "hic", "mic") ) )

# merge wwz and country vars
w1995_2008 %<>% merge(country_vars,
                      by.x = c("year", "Exporting_Country"),
                      by.y = c("year", "ctry")              )

# merge gvc indicators and country vars
gvc_indicators %<>% merge(country_vars, by = c("ctry", "year") )

# merge gvc indicators and nrca
gvc_indicators %<>% merge(nrca, by.x = c("ctry", "isic", "year"), by.y = c("country", "industry", "year") )
```

## General Stylised Facts

 - section 3.1 of wwz
 - growth of GVC trade
 - length of GVC
 - determinants of an industry's participation
 - determinants of a country's participation
 - trade balanses in VA terms differ substantially from those in gross terms

## I2E

  ```{r, echo=FALSE}
 gvc_indicators %>%
  group_by(year, ic) %>%
  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
  ggvis(~year, ~i2e, stroke = ~ic ) %>%
  layer_lines()
 ```

## GVCs of LMICs

 - stylised facts for LMICs
 - source / using countries specific indicators for LMICs
 - nrca and trca of LMIC
 - fvafin vs. fvaint for LMICs (WWZ 3.3)
 - dvix decomposition (WWZ 3.4)

## Conclusion