### 2015-04-22 OECD-TiVa-LMIC-GVCs.Rmd
Updated slieds

---
title: "Global Value Chain Analysis of LMICs"
author: "Victor Kummritz and Bastiaan Quast"
date: "April 22, 2015"
output: ioslides_presentation
---

## Motivation: The rise of GVCs

* "[O]ne of the most important factors shaping contemporary international trade [is] the rise of global value chains."  
*Pascal Lamy, September 2012*
<br />  <br />
* "We value [...] the relevance of regional and global value chains [...] in fostering economic growth, employment and development."  
*G20 leaders, June 2012*
<br /><br />
* "Today, however, production processes increasingly involve global
supply chains spanning multiple countries,[...]."  
*Costinot, Vogel, and Wang, January 2013*

## Motivation: GVCs and development

GVCs offer a new way to development: Countries can join value chains instead of building them:

  + World Economic Forum (2013): Reductions of GVC trade barriers
could raise global GDP by 5\% and international trade by 15\%.
  + UNCTAD (2013): Economies with the fastest growing GVC
participation have GDP per capita growth rates around two percentage
points above the average.

## Motivation: Current shortcomings in the GVCs analysis

 - Limited data availability reduces analysis to high- and middle-income countries, which might not be representative.

 - Analysis so far mostly based on limited set of indicators.

 - Technical complexity of standard GVC indicators represents a large entry barrier to GVC analysis.

## This project

- Develop tools that automate and simplify the derivation of GVC indicators.

- Implement the calculation of novel GVC indicators.

- Analyse these indicators using new data with extensive country coverage.

## New Toolbox and Indicators: decompr and gvc
Background: Leontief decomposition

Applying the Leontief (1936) insight to Inter-Country Input-Output tables to derive value added multipliers ($VB$) following Hummels et al. (1998, 2001):
$$
VB = V + VA + VAA + VAAA + ... = V (I+A+A^{2}+A^{3}+...)
$$
$$
= V (I-A)^{-1}.
$$

The multipliers can be post-multiplied with exports to estimate industry-level value added origins of each industry's exports.

$V$ - Value added / Output, $A$ - Input-output coefficients.

## New Toolbox and Indicators: decompr and gvc
Exemplary indicators:

- I2E (importing-to-export, based on Hummels et al. 1998): The foreign value added content of domestic exports.
    - I2E_LMIC: The foreign value added content sourced in low- and middle income   countries
    - etc.
- E2R (exporting-to-reexport, based on Hummels et al. 1998): The domestic value added content of foreign exports.
- NRCA (New revealed comparative advantage, based on Wang, Wei, and Zhu 2014): Revealed comparative advantage measure based on value added data.

## An application to new data: ICIOS covering LMICs

OECD ICIOs:

- 57 countries

- 34 sectors

- 1995, 2000, 2005, 2008, 2009

- LMICs

## An application to new data: Confirming stylized facts

 ```{r}
 library(decompr)
 ls.str("package:decompr")
 ```

 ```{r}
 library(gvc)
 ls.str("package:gvc")
 ```

 ```{r, echo=FALSE}
 # load the data
load("data/gvc_indicators.RData")
load("data/nrca_df.RData")
load("data/country_vars.RData")
load("data/w1995_2008.RData")
load("data/trca.RData")

 # load required libraries
library(dplyr)
library(ggvis)
library(magrittr)

# contruct avg_pop and gdp
country_vars %<>%
  group_by(ctry) %>%
  summarise(avg_pop = mean(pop)) %>%
  merge(country_vars, . , by = "ctry")

country_vars$avg_gdp <- with(country_vars, avg_pop * avg_gdppc)

# create factor gdp var
country_vars$ic      <- with(country_vars, ifelse(avg_gdppc <= 6000, "lic",
                                                  ifelse(avg_gdppc > 12000, "hic", "mic") ) )

# merge wwz and country vars
w1995_2008 %<>% merge(country_vars,
                      by.x = c("year", "Exporting_Country"),
                      by.y = c("year", "ctry")              )

# merge gvc indicators and country vars
gvc_indicators %<>% merge(country_vars, by = c("ctry", "year") )

# merge gvc indicators and nrca
gvc_indicators %<>% merge(nrca, by.x = c("ctry", "isic", "year"), by.y = c("country", "industry", "year") )
gvc_indicators %<>% merge(trca, by = c("ctry", "isic", "year") )
```

## An application to new data: Confirming stylized facts
The growth of GVC participation.
```{r, echo=FALSE}
gvc_indicators %>%
  group_by(year) %>%
  summarise(e2r = sum(dvar_ik) / sum(exports_ik),
            i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
  ggvis(~year, ~e2r ) %>%
  layer_lines(stroke="e2r") %>%
  layer_lines(~year, ~i2e, stroke="i2e")
```

## An application to new data: Confirming stylized facts
The expansion of GVCs.
```{r, echo=FALSE}
w1995_2008 %>%
  group_by(year) %>%
  summarise( RDV = sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2),
             PDC = sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) %>%
  ggvis(~year, ~RDV, stroke="RDV") %>%
  layer_lines() %>%
  layer_lines(~year, ~PDC, stroke="PDC")
```

## An application to new data: Confirming stylized facts

```{r, echo=FALSE}
gvc_indicators %>%
  group_by(ind_name) %>%
  summarise(i2e = sum(fvax_ams_ik) * 100 / sum(exports_ik) ) %>%
  top_n(10) %>%
  ggvis(~factor(ind_name), ~i2e) %>%
  layer_bars() %>%
  add_axis("x", title = "industry name")
```

## An application to new data: Confirming stylized facts

```{r, echo=FALSE}
gvc_indicators %>%
  group_by(ind_name) %>%
  summarise(e2r = sum(dvar_ik) * 100 / sum(exports_ik) ) %>%
  top_n(10, e2r) %>%
  ggvis(~factor(ind_name), ~e2r) %>%
  layer_bars() %>%
  add_axis("x", title = "industry name")
```

## An application to new data: Confirming stylized facts

```{r, echo=FALSE}
gvc_indicators %>%
  group_by(ctry) %>%
  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
  top_n(10, i2e) %>%
  ggvis(~factor(ctry), ~i2e ) %>%
  layer_bars()
```

## An application to new data: Confirming stylized facts

```{r, echo=FALSE}
gvc_indicators %>%
  group_by(ctry) %>%
  summarise(e2r = sum(dvar_ik) / sum(exports_ik) ) %>%
  top_n(10, e2r) %>%
  ggvis(~factor(ctry), ~e2r ) %>%
  layer_bars()
```

## An application to new data: Confirming stylized facts

```{r, echo=FALSE}
gvc_indicators %>%
  group_by(ctry) %>%
  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik), avg_gdp = mean(avg_gdp) ) %>%
  ggvis(~avg_gdp, ~i2e ) %>%
  layer_points() %>%
  layer_model_predictions(model = "lm")
```

## An application to new data: Confirming stylized facts

```{r, echo=FALSE}
gvc_indicators %>%
  group_by(ctry) %>%
  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik), nat_res_exp_sha_t_3 = mean(nat_res_exp_sha_t_3) ) %>%
  ggvis(~nat_res_exp_sha_t_3, ~i2e ) %>%
  layer_points() %>%
  layer_model_predictions(model = "lm")
```

## An application to new data: Confirming stylized facts

```{r, echo=FALSE}
gvc_indicators %>%
  group_by(ctry) %>%
  summarise(e2r = sum(dvar_ik) / sum(exports_ik), nat_res_exp_sha_t_3 = mean(nat_res_exp_sha_t_3) ) %>%
  ggvis(~nat_res_exp_sha_t_3, ~e2r ) %>%
  layer_points() %>%
  layer_model_predictions(model = "lm")
```

## General Stylised Facts

 - section 3.1 of wwz
 - growth of GVC trade
 - length of GVC
 - determinants of an industry's participation
 - determinants of a country's participation
 - trade balanses in VA terms differ substantially from those in gross terms

## I2E

  ```{r, echo=FALSE}
 gvc_indicators %>%
  group_by(year, ic) %>%
  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
  ggvis(~year, ~i2e, stroke = ~ic ) %>%
  layer_lines()
 ```

## GVCs of LMICs

 - stylised facts for LMICs
 - source / using countries specific indicators for LMICs
 - nrca and trca of LMIC
 - fvafin vs. fvaint for LMICs (WWZ 3.3)
 - dvix decomposition (WWZ 3.4)

## CEQ of China

```{r, echo=FALSE} 
# TRCA vs NRCA for China, ELQ
gvc_indicators %>%
  filter(ctry == "chn") %>%
  filter(ind_name == "elq") %>%
  group_by(year) %>%
  summarise( trca = sum(rca), nrca = sum(nrca) ) %>%
  ggvis(~year, ~trca, stroke = "trca") %>%
  layer_lines() %>%
  layer_lines(~year, ~nrca, stroke="nrca")
```

## Conclusion